---
title: "Homework 7"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
stock <- read.csv("https://bit.ly/3egKiMU")
# Multiplying by 100 to convert to % scale
mydata<-stock*100
```

Problem 1 - Use the Daily stock returns data set. The columns are companies; Man1, Man2, Man3 are manufacturing companies; Serv1, Serv2, Serv3, Serv4 are service companies.


a)	Perform confirmatory factor analysis base on two factors: manufacturing and service.

```{r}
# The p Value = 0.7954455 which is > than 0.05. We fail to reject the null hypothesis. 
# The Data does support the designed CFA model. Model is confirmed.

library(sem)

stock_model <- specifyModel(text = "

Manufacturing -> Man1, lambda1, NA

Manufacturing -> Man2, lambda2, NA

Manufacturing -> Man3, lambda3, NA

ServiceReturn -> Serv1, lambda4, NA

ServiceReturn -> Serv2, lambda5, NA

ServiceReturn -> Serv3, lambda6, NA

ServiceReturn -> Serv4, lambda7, NA

Manufacturing <-> ServiceReturn, rho, NA

Man1 <-> Man1, theta1, NA

Man2 <-> Man2, theta2, NA

Man3 <-> Man3, theta3, NA

Serv1 <-> Serv1, theta4, NA

Serv2 <-> Serv2, theta5, NA

Serv3 <-> Serv3, theta6, NA

Serv4 <-> Serv4, theta7, NA

Manufacturing <-> Manufacturing, NA, 1

ServiceReturn <-> ServiceReturn, NA, 1")

stock_sem <- sem(stock_model, cor(mydata), nrow(mydata))
summary(stock_sem)
```

b)	Report the path diagram that shows coefficient estimates. Code: library(semPlot); semPaths(fitted.sem.object, “est”)

```{r}
library(semPlot)
semPaths(stock_sem, rotation = 2, 'std', 'est')
```

c)	Report SRMR, GFI, and AGFI. What do you conclude? Is the model you made in part (a) approved.

```{r}
# The Data does support the designed CFA model. Model is confirmed.

options(fit.indices = c("GFI", "AGFI", "SRMR"))
criteria = summary(stock_sem)
criteria$SRMR 
criteria$GFI
criteria$AGFI
criteria$SRMR < 0.05
criteria$GFI > 0.95
criteria$AGFI > 0.95
```
d)	Find the 95% confidence interval for the correlation between the two factors: manufacturing and service returns.

```{r}
parameters = summary(stock_sem)
parameters$coeff[8,]$Estimate

conf.L = parameters$coeff[8,]$Estimate - 1.96 * parameters$coeff[8,]$`Std Error`
conf.U = parameters$coeff[8,]$Estimate + 1.96 * parameters$coeff[8,]$`Std Error`
conf.L
conf.U
```

